### 2018-04-12 paper/approach.tex
pass through 4.2.1

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) systems.
%
Solving an CHC system in CDD is coNP time.
%
CDD strictly contains
the union of classes of recursion-free systems that can be solved
in coNP time that have been introduced in previous work.
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all pair of uninterpreted predicates $P, Q \in \predof{S}$ that are siblings(\autoref{sec:chcs}), the 
  transitive dependences of $P$ and $Q$ are disjoint $\tdepsof{P} \cap \tdepsof{Q} = \emptyset$, and there is 
  no uninterpreted predicate shows more than once in a body of clause, then $\cc{S}$ is
  \emph{Clause-Dependence-Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branching and
procedure calls that in each execution path that each statments be
exected at most once.
%
\begin{ex}
  The CHC system $\mcchc$ (given in \autoref{sec:solve-ex}) is a CDD
  system.
  %
  The only clauses in $\mcchc$ that its body has multiple uninterpreted
  predicates are clause (7) that contains $\cc{L9}$ and $\cc{db}$.
  %
  The transitive dependences of $\cc{L9}$ and $\cc{db}$ are \{\cc{L6}\,\cc{L8}\,\cc{L4}\} and $\varnothing$, which are disjoint.
\end{ex}

% talk about body disjoint
The class of CDD systems contains tree structure system, and body-disjoint and linear
systems introduced in previous work.
%
Tree structure system is contained by body-disjoint system. 
%
For recursion-free system $S$, if for each uninterpreted predicate
$Q \in \predof{S}$, there is at most one caluse $C$ that $Q \in \bodyof{C}$,
and there is no uninterpreted predicate shows more than once in a body of clause, then
$S$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
%
If the body of each clause in $\cc{S}$ contains at most
relational predicate, then $\cc{S}$ is
\emph{linear}~\cite{albarghouthi12a}.

% talk about linear system
\begin{thm}
	\label{thm:cdd-contains}
  The class of CDD systems strictly contains the union of the classes
  of body-disjoint and linear systems.
\end{thm}
%
The proof is given in \autoref{app:char}.

@@

% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A CDD System $\cc{S}$.}
  % Output: inductive invariants.
  \Output{If $\cc{S}$ is solvable, then a solution of
    $\cc{S}$; %
    otherwise, the value $\nosoln$.}
  % 
  \myproc{$\solvecdd(\cc{S})$ %
    \label{line:solve-begin}}{ %
    $\sigma \assign \emptyset$ \\
    $\cc{Preds} \assign \topSort(\predof{S})$ \\
    \For {$\cc{P} \in \cc{Preds}$}
    {$\cc{interpolant} \assign \solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$\\
     \Switch{\cc{interpolant}}{
       \lCase{\cc{SAT}:}{\Return{$\nosoln$}}
       \lCase{$\cc{I}$:}{$\sigma$ [$\cc{P} \to \cc{I}$]}
     }
    }
    \Return{$\sigma$}
    }
  %
  \caption{$\solvecdd$: for a CDD system $\cc{S}$, returns a
    solution to $\cc{S}$ or the value $\none$ to denote that
    $\cc{S}$ has no solution.}
  \label{alg:solve-cdd}
\end{algorithm}

% introduce solveCDD
the procedure $\solvecdd$ (\autoref{alg:solve-cdd}) is a solver for CDD systems.
%
$\solvecdd$, given a CDD system $\cc{S}$, it first do a topological sort on all
uninterpreted predicates in $\cc{S}$ based on their dependency relations.
%
In the order of topological sort, for each uninterpreted predicates $\cc{P}$ 
 $\solvecdd$ calles $\solveitp$ to compute the binary interpolation of 
 encoded pre and post formulas with the current solution.
 %
 If the conjunction of pre and post formulas is satisifiable, then $\solvecdd$
 returns $\nosoln$ to indicate this CDD system $\cc{S}$ is not solvable.
 %
 Otherwise, $\solvecdd$ binds the uninterpreted predicate $\cc{P}$ to the
 interpolation result $\cc{I}$ as part of the soltuion.
 %
 The encoding procedure $\prectr$ and $\postctr$ are defined below.
 %
 $\solvecdd$ returns the soltuion $\sigma$.

%
\begin{ex}
  $\solvecdd$, given a CDD system $\mcchc$ in \autoref{sec:running-ex}, may generate
  interpretations queries in any topological ordering of the
  dependency relation of $\mcchc$.
  %
  One such ordering is \cc{L4}, \cc{L6}, \cc{L8}, 
   \cc{L9}, \cc{db} and \cc{main}.
\end{ex}

\subsubsection{Constructing constrains of uninterpreted 
predicate}
%
For a CDD system $S$, given a uninterpreted predicate $\cc{P} \in \predof{S}$, and a partial 
solution $\sigma$ that maps some uninterpreted predicate to its solution formula, 
,the formula $\ctrof{P,\sigma}$ is a compact representation of constrains of $\cc{P}$ based on parital
solution $\sigma$, defined as follows.
% for each relational predicate, define constraint:
If partial solution $\sigma$ does not contains $P$, 
the constraint of $P$ under $\sigma$ that denoted as $\ctrof{P,\sigma}$ 
is constructed as: for all clauses $C \in S$ in which $P$ is the head of $\cc{C}$
\textbf{(1)} each predicate $Q \in \bodyof{C}$ is used and 
\textbf{(2)} the constrained of the clause $\consof{C}$ is encoded.
%
I.e., $\ctrof{P,\sigma}$ is
\[
\biglor_{ (\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
\left( \consof{C_i} \land %
\bigland_{ \cc{Q} \in \bodyof{C_i}} %
b_Q \right)
\]
%
If partial solution $\sigma$ contains $P$, then $\ctrof{P,\sigma}$ is the solution formula $\sigma[P]$.

% introduce the isused predicates:
For each $\cc{P} \in \predof{S}$, there is a corresponding boolean variable
$b_P$.
%
The counterexample characterization of $P$ constrains that
if $P$ is used, which means $b_{P}$ is \cc{True}, then its constraint must hold.
%
I.e., $\vc{P,\sigma}$ is $\neg b_{P} \lor \ctrof{P,\sigma}$.

\begin{ex}
  \label{ex:ctr}
  In the CHC system $\mcchc$ (given in \autoref{sec:solve-ex}), 
  when $\mcsolve$ solves the uninterpreted predicate \cc{L9}, it 
  has a partial solution $\sigma$ that maps \cc{L6} to $n \ge 0$.
  %
  For predicate \cc{L6}, the the $\ctrof{L6,\sigma}$ is $n \ge 0$.
  %
  For the 
  uninterpreted \cc{L9}, the $\ctrof{L9,\sigma}$ is $(abs' = n \land b_{L6}) \lor (abs' = n \land b_{L8})$.
  %
  The $\vc{L9,\sigma}$ is $\neg b_{L9} \lor \ctrof{L9,\sigma}$.
\end{ex}

>>>>

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) systems. As discussed in
\autoref{sec:overview}, CDD systems \textbf{(1)} are a superclass of other
practical classes of directly solvable CHC systems and \textbf{(2)} can be
solved in co-NP.
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all sibling pairs, $P, Q \in \predof{S}$,
  %
  the transitive dependences of $P$ and $Q$ are disjoint ($\tdepsof{P} \cap
  \tdepsof{Q} = \emptyset$)
  %
  and no predicate is a sibling of itself,
  %
  then $\cc{S}$ is \emph{Clause-Dependence Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branches and procedure calls such
that each execution path invokes each statement at most once.
%
\begin{ex}
  The CHC system $\mcchc$ is a CDD system. An argument is given in
  \autoref{sec:solve-ex}.
\end{ex}

% talk about previous system
The class of CDD systems contains body-disjoint and linear
systems introduced in previous work.
%
For recursion-free system $S$, if for each uninterpreted predicate
$Q \in \predof{S}$, there is at most one caluse $C$ that $Q \in \bodyof{C}$,
and there is no uninterpreted predicate shows more than once in a body of clause, then
$S$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
%
If the body of each clause in $\cc{S}$ contains at most
relational predicate, then $\cc{S}$ is
\emph{linear}~\cite{albarghouthi12a}.

% talk about linear system
\begin{thm}
	\label{thm:cdd-contains}
  The class of CDD systems strictly contains the union of the classes
  of body-disjoint and linear systems.
\end{thm}
%
Proof given in \autoref{app:char}.

@@

% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A CDD System $\cc{S}$.}
  % Output: inductive invariants.
  \Output{If $\cc{S}$ is solvable, then a solution of
    $\cc{S}$; %
    otherwise, the value $\nosoln$.}
  % 
  \myproc{$\solvecdd(\cc{S})$ %
    \label{line:solve-begin}}{ %
    $\sigma \assign \emptyset$ \\
    $\cc{Preds} \assign \topSort(\predof{S})$ \\
    \For {$\cc{P} \in \cc{Preds}$}
    {$\cc{interpolant} \assign \solveitp(\prectr(\cc{P},\sigma),\postctr(\cc{P},\sigma))$\\
     \Switch{\cc{interpolant}}{
       \lCase{\cc{SAT}:}{\Return{$\nosoln$}}
       \lCase{$\cc{I}$:}{$\sigma$ [$\cc{P}$] $\assign$ $\cc{I}$}
     }
    }
    \Return{$\sigma$}
    }
  %
  \caption{$\solvecdd$: for a CDD system $\cc{S}$, returns a
    solution to $\cc{S}$ or the value $\none$ to denote that
    $\cc{S}$ has no solution.}
  \label{alg:solve-cdd}
\end{algorithm}

% introduce solveCDD
\autoref{alg:solve-cdd} presents $\solvecdd$, a procedure designed to solve
CDD systems.
%
Given a CDD system $\cc{S}$, $\solvecdd$ topologically sorts the uninterpreted
predicates in $\cc{S}$ based on their dependency relations.
%
Then, the algorithm calculates interpretations for each predicate in this order
by invoking $\solveitp$.
%
$\solveitp$ computes a binary interpolant of the pre and post formulas of the
given predicate, where these formulas are based on the current, partial
solution.
%
The pre and post formulas are computed respectively by $\prectr$ and
$\postctr$, which we define below.
%
It is possible that the pre and post formulas may be mutually satisifiable, in
which case $\solveitp$ returns \cc{SAT}. In this case, $\solvecdd$ returns
$\nosoln$ to indicate that $\cc{S}$ is not solvable.
%
Otherwise, $\solvecdd$ modifies the partial solution by setting the
interpretation of $\cc{P}$ to $\cc{I}$.
%
Once all predicates have been interpolated, $\solvecdd$ returns the complete
solution, $\sigma$.

%
\begin{ex}
  Given the CDD system $\mcchc$, $\solvecdd$ may generate interpolation
  queries in any topological ordering of the dependency relations.
  %
  One such ordering is $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$, $\cc{L}_9$, \cc{db},
  \cc{main}.
\end{ex}

\subsubsection{Constructing constraints for predicates}
%
For a CDD system $S$, a predicate $\cc{P} \in \predof{S}$, and a partial
solution $\sigma$ that maps predicates to their solutions, the formula
$\ctrof{P,\sigma}$ is a compact representation of the constraints of $\cc{P}$.
%
If $\sigma$ does not contain $P$, then the constraint of $P$, denoted
$\ctrof{P,\sigma}$, is constructed from the clauses where $P$.  When $\sigma$
does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$:
%
\[
  \ctrof{P,\sigma}=
  \begin{dcases}
    \biglor_{(\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
    \left( \consof{C_i} \land %
      \bigland_{ \cc{Q} \in \bodyof{C_i}} \cc{b}_Q
    \right),
  &\text{if } P \notin \sigma\\
  \sigma[P], &\text{if } P \in \sigma
  \end{dcases}
\]
%
% introduce the isused predicates:
Each $\cc{P} \in \predof{S}$ has a corresponding boolean variable $b_P$.
%
The counterexample characterization of $P$ constrains that if $P$ is used,
(meaning $\cc{b}_{P} = True$) then its constraint must hold:
%
$\vc{P,\sigma} = \neg \cc{b}_{P} \lor \ctrof{P,\sigma}$.

\begin{ex}
  \label{ex:ctr}
  When $\mcsolve$ solves the predicate $\cc{L}_9$ in $\mcchc$, it generates a
  constraint based on clauses (5) and (6):
  $\ctrof{\cc{L}_9,\sigma} =
    (\cc{abs'} = \cc{n} \land \cc{b}_{\cc{L}_6})
    \lor
    (\cc{abs'} = -\cc{n} \land \cc{b}_{\cc{L}_8})$.
  The counterexample characterization for $\cc{L}_9$ is based on its boolean
  indicator and this constraint:
  $\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$.
\end{ex}
